---
title: 'Assignment #5'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(countrycode)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
```


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
coordinates <- cbind(world, st_coordinates(st_centroid(world$geometry)))
coordinates <- 
  coordinates %>%
  select(name,sovereignt, type, X,Y)
```



## Exercises

(@) Convert the name of a country of choice into five different languages of your choice. Use the world dataset and use countryname. 

```{r}
question1 <-
  world %>%
  filter(name == 'Argentina')

countryname(question1$name, 'cldr.name.zh')
countryname(question1$name, 'cldr.name.es')
countryname(question1$name, 'cldr.name.ru')
countryname(question1$name, 'cldr.name.ar')
countryname(question1$name, 'cldr.name.fr')
```


(@) Find the ‘region’ name and code for the top 30 countries. Use ‘countrycode’ for this.

```{r}
question2 <-
  world %>% 
  head(30)

countrycode(question2$name, origin = 'country.name', destination = 'region')
```


(@) Use the guess_field() function to find the closest match of the origin of the adm0_a3 column in the world dataset.

```{r}
guess_field(world$adm0_a3)
```


(@) Draw a map with 5 countries with different labels using geom_sf, geom_text and countryname

```{r}
question4 <-
  coordinates %>% 
  filter(name %in% c("United States", "Canada", "Greenland", "Mexico", "Cuba")) %>% 
  mutate(label = countryname(name, destination = 'un.name.en'))

ggplot() + 
  geom_sf(data = question4) + 
  geom_text(data=question4 ,aes(X, Y, label=name), size=3) +
  labs(x = NULL,
       y = NULL) + 
  theme(panel.background = element_blank())
```


(@) Draw a map that has each country named after the continent it belongs to and colored by their continent.

```{r}
question5 <-
  coordinates %>% 
  mutate(label = countryname(name, destination = 'region'))

ggplot(data = world) +
    geom_sf(data = question5, aes(fill = label)) + 
    geom_text(data=question5 ,aes(X, Y, label=world$continent),
    color = "darkblue", check_overlap = TRUE, size=2)+
    labs(x = NULL,
         y = NULL,
         fill = "Region",
         title = "World Map with Continent Labels, Colored by Region")+
    theme(panel.background = element_blank())
```






