---
title: "Assignment #3"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
# SEE modeldata package for new datasets
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(stacks)            # for stacking models
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(DALEX)             # for model interpretation  
library(DALEXtra)          # for extension of DALEX
library(patchwork)         # for combining plots nicely
library(dbplyr)            # for SQL query "cheating" - part of tidyverse but needs to be loaded separately
library(mdsr)              # for accessing some databases - goes with Modern Data Science with R textbook
library(RMySQL)            # for accessing MySQL databases
library(RSQLite)           # for accessing SQLite databases

#mapping
library(maps)              # for built-in maps
library(sf)                # for making maps using geom_sf
library(ggthemes)          # Lisa added - I like theme_map() for maps :)

#tidytext
library(tidytext)          # for text analysis, the tidy way!
library(textdata)          
library(reshape2)
library(wordcloud)         # for wordcloud
library(stopwords)
```

## Put it on GitHub!

[Here](https://github.com/alexdenzler/STAT494_site_Denzler) is my GitHub link.


## Local Interpretable Machine Learning

**Tasks**

  1. Choose 3 new observations and do the following for each observation:
  * Construct a break-down plot using the default ordering. Interpret the resulting graph. Which variables contribute most to each observation’s prediction?
  * Construct a SHAP graph and interpret it. Does it tell a similar story to the break-down plot?
  * Construct a LIME graph (follow my code carefully). How close is each original prediction to the prediction from the local model? Interpret the result. You can also try using fewer or more variables in the local model than I used in the example.
  
  2. Describe how you would use the interpretable machine learning tools we’ve learned (both local and global) in future machine learning projects? How does each of them help you?
  

## SQL 

You will use the `airlines` data from the SQL database that Lisa used in the example in the tutorial. Be sure to include the chunk to connect to the database here. And, when you are finished, disconnect. You may need to reconnect throughout as it times out after a while.
  
  







