---
title: "Assignment #4"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
##knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
#Regular expressions
library(tidyverse)        # contains stringr for regex
library(googlesheets4)    # for reading in data
gs4_deauth()              # to read in google sheet (or download)

#tmap
library(tmap)
library(pacman)
```


## Put it on GitHub

[Here](https://github.com/alexdenzler/STAT494_site_Denzler) is my GitHub link.

## Regular Expressions

```{r cache=TRUE}
bestsellers <- read_sheet("https://docs.google.com/spreadsheets/d/1n3xKHK4-t5S73LgxOJVJWT5fMYjLj7kqmYl1LHkpk80/edit?usp=sharing")
```

  1. Find books with multiple authors (HINT: Consider the possibility of an author having “and” in their name)
  
```{r}
bestsellers %>% 
  mutate(mult_auth = str_detect(author, "\\sand\\s")) %>% 
  filter(mult_auth == "TRUE") %>% 
  summarize(title, author, mult_auth)
```
  
  
  2. Detect if the author’s first name starts with a vowel
  
```{r}
bestsellers %>% 
  mutate(vowel = str_detect(author, "^[AEIOU]")) %>% 
  filter(vowel == "TRUE") %>% 
  summarize(author, vowel)
```
  
  3. Change all authors with the name Alexander to Alex
  
```{r}
bestsellers %>% 
  mutate(names = str_replace_all(author, "Alexander", "Alex"))
```
  
  4. Find books that are the second book in a series

```{r}
bestsellers %>%
  mutate(series = str_detect(bestsellers %>% pull(description), 
                             pattern = "second book")) %>%
  filter(series == TRUE)
```
  
  
  5. Find books that are the third or fourth one in a series
  
```{r}
bestsellers %>%
  mutate(series = str_detect(bestsellers %>% pull(description), 
                          pattern = "(third|fourth) book")) %>%
  filter(series == TRUE)
```
  
  6. Find books that are the 10th, 11th, …, or 19th book in a series
  
```{r}
bestsellers %>%
  mutate(series = str_detect(bestsellers %>% pull(description), 
                          pattern = "1[0-9]th book")) %>%
  filter(series == TRUE)
```
  
  
  7. Describe in your own words how you would go about writing a regular expression for password pattern matching (i.e. 8 character minimum, one capital letter minimum, one lowercase letter minimum, one digit minimum, one special character minimum)
  
  * Each regular expression would have to check to see if the range or requirements were included in the password. We would be able to use `str_detect()` and have each regular expression check for a capital letter, a lowercase letter, a digit, and a special character (whcih could be written like this $[a-z][A-Z]\d\W$). 
  
  
## `tmap` Exercises


  
  
  
  


